% kate: word-wrap true;

% IMPORTANT NOTE: I've never dealt much with syntax before, at least not in a
% way as formal as I'm trying to come to terms with it here. If there's
% anything wrong with my analyses, it's because I don't know enough yet. Things
% may be subject to change *especially* in this chapter.

\chapter{Phrase structures}

While the previous chapter dealt largely with the various parts of speech and
their various distributive and inflectional properties, the present chapter
will elaborate on how these words combine into syntactic phrases. Since Ayeri
is a verb-initial language, it is probably most comfortably analyzed in terms
of Lexical-Functional Grammar (\cite{bresnan1982}~ff.; more recently
\cite{bresnan2016}), since \Lfg{} does not require complicated derivations
behind the surface structure of  sentences.\footnote{Passivization, for
instance, is assumed to be a lexically  motivated alternation in predicate
structure (\Sbj{} is blocked, so the  nominative is assigned to \Obj{}, and the
original \Sbj{} is expressed  by an \Adjc{}), rather than an internal
derivational process \citep[23\psqq]{bresnan2016}.} It will be assumed here
that, even though Ayeri is basically VSO with predicate and predication not
adjacent to each other, it is configurational in that there is a VP which
c-commands a number of other constituents as complements in transitive
sentences.

In principle, \Lfg{} assumes that grammar operates on different structural 
levels: mainly, these are a(rgument) structure, c(onstituent) structure, and 
f(unctional) structure; other layers have been proposed by different 
researchers for different purposes \citep[862--865]{buttking2015}. 
\citet{bresnan2016} define three core design principles for \Lfg{}:

\begin{description}
\item[Variability:] \textcquote[41]{bresnan2016}{The principle of variability
states that \emph{external structures vary across languages}. The formal model
of external structure in \Lfg{} is the \emph{c-structure}, which stands for
\enquote{constituent structure} or \enquote{categorial structure}}. 
C-structures are commonly represented by context-free phrase-structure rules; 
constituency trees are based on an extended version of X-bar theory 
\citep[42]{bresnan2016}.\footnote{The basic recursive rules of X-bar theory 
are observed:
\begin{enumerate}[nosep, leftmargin={2\footnotemargin}]
\item XP → YP, \xbar{X} (specifier rule)
\item \xbar{X} → \xbar{X}, YP (adjunct rule)
\item \xbar{X} → \xhead{X}, YP (complement rule)
\end{enumerate}

The principle of economy of expression furthermore dictates that essentially, 
trees be pruned of empty terminal nodes and non-branching preterminal nodes, 
since these do not provide structurally or semantically relevant information 
\citep[119--128]{bresnan2016}.}

\item[Universality:] \textcquote[42]{bresnan2016}{The principle of universality
states that \emph{internal structures are largely invariant across languages}.
The formal model of internal structure in \Lfg{} is the \emph{f-structure},
which stands for \enquote{functional structure}}. The f-structure is depicted
as an argument-value matrix (\Avm{}) which maps the relations between `subject'
(\Sbj{}), `object' (\Obj{}), `predicator' (\Pred{}), etc. as functional
abstractions of NP, VP, V, etc. \citep[42]{bresnan2016}. Verbs are also
presented with their \fw{a-structure} spelled out. That is, which arguments a
verb has relations to is formally stated \citep[15]{bresnan2016}. The
f-structure collates semantic features associated with heads of grammatical
functions (GFs), such as case (\Case{}), person (\Pers{}), number (\Num{}),
which are abstract features and as such need not have morphological realization
\citep[43]{bresnan2016}.

\item[Monotonicity:] \textcquote[43]{bresnan2016}{Constituent structure form is
simply not the same in all languages [...] In \Lfg{} the correspondence mapping
between internal and external structures does not preserve sameness of form.
Instead, \emph{it is designed to preserve inclusion relations between the
information expressed by the external structure and the content of the internal
structure}}. Due to the principle of monorepresentation, information
distributed over different morphemes which logically belongs to a single
grammatical function is presented in the f-structure as unified.

\end{description}

\begin{figure}[t]\centering
\caption[F-structure mappings]{F-structure mappings \citep[15]{bresnan2016}}

\begin{tabular}[t]{l @{\quad\quad} c}
argument (a-)structure:
& \astruct{\tikzmark{verb}verb}{\tikzmark{x}x, \tikzmark{y}y}\bigskip \\

functional (f-)strucutre:
& \tikzmark{fstruct}{\smaller\begin{avm}
\[
	\quad sbj \tikzmark{subj} & \[
		{\enspace}\vdots{\enspace} \\
	\]{\quad} \\
	
	\quad obj \tikzmark{obj} & \[
		{\enspace}\vdots{\enspace} \\
	\] \tikzmark{objval}{\quad} \\
	\quad \tikzmark{pred} pred & \dots \tikzmark{predval} \\
\]
\end{avm}}\bigskip\\

constituent (c-)structure:
& \begin{forest} baseline
[\xbar{V}
	[\subnode{V}{V}]
	[NP \tikzmark{NP}
		[\xbar{N} \tikzmark{Nbar}]
	]
]
\end{forest}
%
\end{tabular}
\begin{tikzpicture}[remember picture, overlay]
\draw [-latex]
	([xshift=1.5ex, yshift=-0.5ex]{pic cs:verb})
	to [out=south, in=north west]
	([yshift=1ex]{pic cs:pred});
	
\draw [-latex]
	([xshift=0.5ex, yshift=-0.5ex]{pic cs:x})
	to [out=south, in=north east]
	([yshift=1ex]{pic cs:subj});
	
\draw [-latex]
	([xshift=0.5ex, yshift=-0.75ex]{pic cs:y})
	to [out=south, in=north east]
	([yshift=1ex]{pic cs:obj});
	
\draw [-latex]
	([yshift=0.5ex]{pic cs:V})
	to [out=west, in=south west]
	([yshift=-10ex]{pic cs:fstruct});
	
\draw [-latex]
	([yshift=0.5ex]{pic cs:NP})
	to [out=east, in=east]
	([yshift=0.5ex]{pic cs:objval});
	
\draw [-latex]
	([yshift=0.5ex]{pic cs:Nbar})
	to [out=east, in=east]
	([yshift=0.5ex]{pic cs:objval});
\end{tikzpicture}
\label{fig:phimap}
\end{figure}

To illustrate the different parallel structures in operation,
\citet[15]{bresnan2016} give the schema in \autoref{fig:phimap} to demonstrate
which part of the a- and c-structure respectively corresponds (`links', `maps')
to which part of the f-structure.
% \footnote{\citet{bresnan2016} use \textsc{`subj'} for `subject'. ; for
% consistency with the above I will use `\Sbj{}' in the following. I will also
% divergently use \Compl{} and \XCompl{} for \textsc{(x)comp}, since
% \textsc{comp} has already been used for `comparative' above.}
Regarding the different functions distinguished, \Lfg{} 
assumes the following hierarchies \citep[97, 100]{bresnan2016}:

\pex\label{ex:functions}
\a\label{ex:gfs} Grammatical functions (GFs):\\
	$\overbrace{\Sbj{} > \Obj{} > \SObj{}}^{\text{core}} > 
	\overbrace{\OblqT{} > \XCompl{}, \Compl{} > \Adjc{}}^{\text{noncore}}$
\a\label{ex:nonafs} (Non)argument functions (AFs/$\overline{\mbox{AF}}$s):\\
	$\underbrace{\Top{}\: \Foc{}}_{\text{non-a-fns}}\; 
	\overbrace{\Sbj{}\: \Obj{}\: \SObj{}\: \OblqT{}\: \XCompl{}\: 
		\Compl{}}^{\text{a-fns}}\; 
	\underbrace{\Adjc{}}_{\text{non-a-fns}}$
\a\label{ex:dfs} Discourse functions (DFs):\\
	$\overbrace{\Top{}\: \Foc{}\: \Sbj{}}^{\text{d-fns}}\;  
	\underbrace{\Obj{}\: \SObj{}\: \OblqT{}\: \XCompl{}\: \Compl{}\: 
		\Adjc{}}_{\text{non-d-fns}}$
\xe

The elements listed in (\ref{ex:functions}) will also appear in
phrase-structure rules and c-structure trees together with arrows. These arrows
symbolize inheritance of feature information from the current level (↓) of the
tree to the next (↑), so for instance, `\pass{\Sbj}' means that the information
subsumed by the current node (`down') is passed on as the subject function of
the next higher node (`up') in the tree. Concise information on notational
formalisms of \Lfg{} can be found, for instance, in \citet{buttking2015}.

\section{Noun phrases and determiner phrases}
\label{sec:nps-dps}

Noun phrases (NPs), and determiner phrases (DPs) as their functional 
counterpart, fulfill the functions of subject (\Sbj{}), object (\Obj{}), 
secondary object (\SObj{}), as well as various oblique constituents (\OblqT;
see (\ref{ex:dpnpcasemap}) below), and they can also form adjuncts (\Adjc{}).
DPs and NPs can also constitute topics (\Top{}). Which DP or NP receives which 
function is selected by the a-structure of the verb---this also has 
repercussions on case- and topic marking.

Since the constituent containing the various non-verbal elements in Ayeri is 
likely exocentric and constituents may move around within it in a restricted 
way, we have to assume that not constituent structure, but case marking 
identifies the grammatical functions of the various arguments of verbs. Thus, 
the following lexicocentric conditions operate on both DP and NP as exponents 
of case:

\ex\label{ex:dpnpcasemap}\labels
\begin{tabular}[t]{@{} l @{\quad} l @{$\implies$} l}
\tl\quad & \downs{\Case} = \Aarg	& \pass{\Sbj} \\
\tl\quad & \downs{\Case} = \Parg	& \pass{\Obj} \\
\tl\quad & \downs{\Case} = \Dat		& \pass{\SObj} \logor{}
										\pass{\Oblq{loc}} \\
\tl\quad & \downs{\Case} = \Gen		& \pass{\Oblq{poss}} \logor{} 
						\pass{\Oblq{loc}} \\
\tl\quad & \downs{\Case} = \Loc		& \pass{\Oblq{loc}} \\
\tl\quad & \downs{\Case} = \Caus	& \pass{\Oblq{caus}} \\
\tl\quad & \downs{\Case} = \Ins		& \pass{\Oblq{ins}} \logor{} 
										\pass{\Adjc} \\
\end{tabular}
\xe

The rules in (\ref{ex:dpnpcasemap}) determine the typical mappings between case
marking and grammatical functions, which are not always unambiguous. As
explained above (compare \autoref{subsec:case}), the dative case does not only
indicate that something is done to this referent or to their benefit, but it
may also indicate motion towards this referent. Likewise, the genitive case
does not only indicate possession, but also origin, and motion from this
referent. Nominal adjuncts to nouns which specify what the noun consists also
appear in the instrumental case, besides the instrumental being used to
indicate the means or the circumstance by which an action comes about.
Moreover, DPs or NPs may also lack case marking, which indicates that the
respective phrase is a part of the topic of the verb, which is what
(\ref{ex:topicrule}) describes:

\ex\label{ex:topicrule}
¬\,\downs{\Case} $\implies$ \elem{\Top}
\xe

Instead of case marking on the DP or NP, there is a marker in front of the verb
which provides information on the case and, if \AgtT{} or \PatT{}, also about
the animacy of the topicalized phrase. This means that grammatical information
about the topic of a phrase is spread over two discontinuous locations. This
issue does not pose a problem to an \Lfg{}-based analysis, however, since both
locations unify their information content in the f-structure feature \Top{}.
Since information located in multiple places is jointly feeding this feature, I
am using the annotation `\elem{\Top}' for each location rather than simple
`\pass{\Top}'. Note that only one NP among the arguments of a verb may be the
topic of the phrase, and a topic can only be marked if the verb is finite and
the number of arguments to the verb is greater than one.

\subsection{Noun phrases}

Nouns are one of the main parts of speech of Ayeri, and nouns can be modified
by a number of other free elements, as we have seen previously---adjectives,
possessive adjectives, as well as relative clauses and nominal adjuncts. These
typically follow nouns. It was also described before that Ayeri's nouns may
host a number of clitics, among which are deictic prefixes and quantifiers, as
well as enclitic case markers in the case of proper nouns. These clitics,
however, will not be treated as targets of syntactic operations, since \Lfg{}
follows the approach of lexical integrity. Thus, bound elements like affixes
and clitics are assumed not to be reflected or affected by syntax itself. The
phrase structure of NPs should thus look like depicted in (\ref{ex:npstruct}).

\pex\label{ex:npstruct}
\a NP → \anno*{\xbar{N}}
\a \xbar{N} → \anno*{\xhead{N}} $\left(\anno*[{\pass{\Adjc}}]{XP}\right)$
\xe

This rule defines that NPs have a lexical head which is on the left side,
followed optionally by modifiers which act as adjuncts to the noun. The
description leaves open which kind of phrases the adjuncts can consist of, but
as mentioned above, these are commonly adjectives (forming adjective phrases,
APs), nominal adjuncts (NPs), as well as relative clauses (forming
complementizer phrases, CPs). This can be represented as a constituent-structure tree in the way described in (\ref{ex:npcstruct}).

\ex\label{ex:npcstruct}\labels
\begin{forest}
[{\anno[\{\pass{df} | \pass{gf}\}]{NP}}
	[\anno{\xbar{N}}
		[\anno{\xhead{N}}]
		[{$\left(\anno[{%
				\pass{\Adjc}%
			}]{XP}\right)$
		}]
	]
]
\end{forest}
\xe

Here as well, we can see a nominal head on the left which may be modified by
adjuncts of various types. The maximal projection of \xhead{N} (that is, NP) is
annotated very generally for the function of the NP---basically, an NP can act
as either a discourse function (DF) or a grammatical function (GF). 
(\ref{ex:nounmods}) gives an example of each kind of modifier. Since there is 
no grammatical context given, NP is unmarked for function in these examples.

\pex\label{ex:nounmods}
\a %
	\begin{minipage}[t]{.5\linewidth}
	\begingl
		\glpreamble noun + adjective: //
		\gla ningan hiro //
		\glb ningan hiro //
		\glc story new //
		\glft `new story' //
	\endgl
	\end{minipage}
	~
	\begin{forest} shorter edges,
	[NP
		[\anno{\xbar{N}}
			[\anno{\xhead{N}}
				[{ningan}]
			]
			[{\anno[\pass{\Adjc}]{AP}}
				[{hiro}, roof]
			]
		]
	]
	\end{forest}

\a %
	\begin{minipage}[t]{.5\linewidth}
	\begingl
		\glpreamble noun + genitive attribute: //
		\gla kegan ayonena //
		\glb kegan ayon-ena //
		\glc hat man-\Gen{} //
		\glft `the man's hat' //
	\endgl
	\end{minipage}
	~
	\begin{forest} shorter edges,
	[NP
		[\anno{\xbar{N}}
			[\anno{\xhead{N}}
				[{kegan}]
			]
			[{\anno[\pass{\Adjc}]{NP}}
				[{ayonena}, roof]
			]
		]
	]
	\end{forest}

\a %
	\begin{minipage}[t]{.5\linewidth}
	\begingl
		\glpreamble noun + instrumental attribute: //
		\gla kasu bariri //
		\glb kasu bari-ri //
		\glc basket meat-\Ins{} //
		\glft `basket of meat' //
	\endgl
	\end{minipage}
	~
	\begin{forest} shorter edges,
	[NP
		[\anno{\xbar{N}}
			[\anno{\xhead{N}}
				[{kasu}]
			]
			[{\anno[\pass{\Adjc}]{NP}}
				[{bariri}, roof]
			]
		]
	]
	\end{forest}

\a %
	\begin{minipage}[t]{.5\linewidth}
	\begingl
		\glpreamble noun + relative clause: //
		\gla nanga si incāng //
		\glb nanga si int=yāng //
		\glc house \Rel{} buy=\TsgM{}.\Aarg{} //
		\glft `the house he bought' //
	\endgl
	\end{minipage}
	~
	\begin{forest} shorter edges,
	[NP
		[\anno{\xbar{N}}
			[\anno{\xhead{N}}
				[{nanga}]
			]
			[{\anno[\pass{\Adjc}]{CP}}
				[{si incāng}, roof]
			]
		]
	]
	\end{forest}

\xe

As described before (compare \autoref{subsec:clitics}), nouns can be modified
by a number of clitics which are not represented through syntax. Since it is
not possible for these clitic elements to be divided from their phonological
hosts, they should be treated as being an integral part of the word they attach
to. Hence, \xhead{N} is given in (\ref{ex:nouncltree}) as split into `Cl' and
\xhead{N}.

\pex\label{ex:nouncltree}
\a %
	\begin{minipage}[t]{.5\linewidth}
	\begingl
		\glpreamble noun + deictic prefix: //
		\gla eda- @ nanga //
		\glb eda= nanga //
		\glc this= house //
		\glft `this house' //
	\endgl
	\end{minipage}
	~
	\begin{forest} shorter edges,
	[\anno{\xhead{N}}
		[\anno{Cl}
			[eda-]
		]
		[\anno{\xhead{N}}
			[nanga]
		]
	]
	\end{forest}

\a %
	\begin{minipage}[t]{.5\linewidth}
	\begingl
		\glpreamble noun + quantifier: //
		\gla nangās @ -ikan //
		\glb nanga-as =ikan //
		\glc house-\Parg{} =many //
		\glft `many houses' //
	\endgl
	\end{minipage}
	~
	\begin{forest} shorter edges,
	[\anno{\xhead{N}}
		[\anno{\xhead{N}}
			[nangās]
		]
		[\anno{Cl}
			[-ikan]
		]
	]
	\end{forest}

\a %
	\begin{minipage}[t]{.5\linewidth}
	\begingl
		\glpreamble proper noun + case: //
		\gla ang @ Diyan //
		\glb ang= Diyan //
		\glc \Aarg{}= Diyan //
		\glft `Diyan' //
	\endgl
	\end{minipage}
	~
	\begin{forest} shorter edges,
	[\anno{\xhead{N}}
		[\anno{Cl}
			[ang]
		]
		[\anno{\xhead{N}}
			[Diyan]
		]
	]
	\end{forest}

\xe

Of course, it is also possible to combine these nominal modifiers. In this
case, there is a certain hierarchy, presumably based on Behaghel's first law,
\textquote{Das oberste Gesetz ist dieses, daß das geistig eng Zusammengehörige 
auch eng zusammengestellt wird} (\cite[4]{behaghel1932}; `The supreme law is
such that the mentally closely related is also arranged in close proximity.'), 
and also grammatical weight \citep{wasow1997}:

\begin{enumerate}[noitemsep]
	\item instrumental NP indicating what the head consists of,
	\item APs (also cardinals) and other NPs describing attributes,
	\item possessive genitive NPs,
	\item relative clauses.
\end{enumerate}

\citet{wasow1997} writes that \textcquote[102]{wasow1997}{[i]t is very hard to
distinguish among various structural weight measures as predictors of weight
effects. Counting words, nodes, or phrasal nodes all work well}, which means
that no single metric can be used to describe the order of constituents in a
phrase. However, for instance, relative clauses trail whenever possible
presumably since they tend to contain whole subclauses and thus a lot of
information. It seems advisable not to put an element with much less 
information content after them, especially when it refers to a different head
than all the things inside the relative clause. The following example 
(\ref{ex:nounmodord}) illustrates the unmarked order of modifiers.

\pex\label{ex:nounmodord}
\a\begingl
	\gla diranang caban nā si ang mica ya @ Kārvisam //
	\glb diran-ang caban nā si ang mit=ya.Ø ya= Kārvisam //
	\glc uncle-\Aarg{} favorite \Fsg{}.\Gen{} \Rel{} \AgtT{} 
		live=\TsgM{}.\Top{} \Loc{}= Kārvisam //
	\glft `my favorite uncle who lives in Kārvisam' //
\endgl
\medskip

\a\begin{forest} shorter edges, narrower nodes,
[{\anno[\pass{\Sbj}]{NP}}
	[\anno{\xbar{N}}
		[\anno{\xbar{N}}
			[\anno{\xbar{N}}
				[\anno{\xhead{N}}
					[{diranang}]
				]
				[{\anno[\elem{\Adj}]{AP}}
					[{caban}, roof]
				]
			]
			[{\anno[\elem{\Adj}]{AP}}
				[{nā}, roof]
			]
		]
		[{\anno[\elem{\Adj}]{CP}}
			[{si ang mica ...}, roof]
		]
	]
]
\end{forest}

\xe

As the c-structure tree in (\ref{ex:nounmodord}) shows, Ayeri prefers
head--dependent word order with exceeding consistency. As illustrated by
previous examples, both adjuncts and complements are, for the most part,
consistently appended to the right of their heads, which means that Ayeri may
be classified as a rather consistently right-branching language. However, a
certain number of postpositions form an exception to this classification
(\autoref{subsec:postpos}). In the light of word order typology, we can
formulate the following generalizations:

\pex
\a Order of noun and adjective: N Adj
\a Order of noun and genitive: N Gen
\a Order of noun and relative clause: N Rel
\xe

More important to \Lfg{} than c-structure trees, however, are
function-structure matrices which gather all information in a given utterance
and gather potentially disparate information into semantically coherent units
within the matrix.\footnote{Essentially, c-structure is similar to the tree
hierarchy of paragraphs, images, tables etc. in an \textsc{html} file, while
f-structure describes semantic properties of elements in the tree similar to
how \textsc{css} defines the layout properties of these elements.} In the
following, I will thus give a list of morpholexic specifications which give an
overview of the different semantic and morphological features nouns basically
provide (also compare \autoref{sec:nouns}). These also form the basis for
f-structure matrices of the kind already shown in (\ref{ex:clitics_43}),
\autoref{clitics_postverb_person}, p.~\pageref{ex:clitics_43}.

\ex\label{ex:nounmorphlex}%
\adjustbox{valign=t}{%
	\begin{tabu} {\usetabu{morphlex}}
	...
		& N
		& \begin{tabular}[t]{l l l}
			\ups{\Pred} & = & `...' \\
			\ups{\Anim} & = & $\pm$ \\
			\ups{\Case} & = & \{\Aarg{}, \Parg{}, \Dat{}, \Gen{}, 
				\Loc{}, \Ins{}, \Caus{}\} \\
			\ups{\Gend} & = & \{\M{}, \F{}, \N{}, \Inan{}\} \\
			\ups{\Num} & = & \{\Sg{}, \Pl{}\} \\
			\ups{\Pers} & = & 3 \\
		\end{tabular}
	\end{tabu}%
}
\xe

Nouns generally imply a third-person reference; they distinguish number, gender
and animacy, as well as case. Clitics, however, may also add information about
deixis (\ref{ex:deixisfeat}); likeness and quantity might be interpreted
conveniently as adding to the list of a noun's \Adjc{} feature 
(\ref{ex:nounqadj}).

\ex\label{ex:deixisfeat}
%\adjustbox{valign=t}{%
	\begin{tabu} {\usetabu{morphlex}}
	\hphantom{...}
		& \hphantom{N}
		& \begin{tabular}[t]{l l l}
			\ups{\Deix} & = & \{$this$, $that$, $such$\} \\
		\end{tabular}
	\end{tabu}%
%}
\xe

\pex~\label{ex:nounqadj}
\a\begingl
	\gla ganang-hen mino //
	\glb gan-ang=hen mino //
	\glc child-\Aarg{}=all happy //
	\glft `all happy children' //
\endgl

\a\begin{avm}
\[
	\Subj{}	&	\[
					\Pred	&	`child' \\
					\Anim	&	$+$ \\
					\Case	&	\Aarg \\
					\Adjc	&	\{
									\[
										\Pred	&	`all' \\
										\Num	&	\Pl \\
									\], \\
									\[
										\Pred	&	`happy' \\
									\] \\
								\} \\
				\] \\
\]
\end{avm}
\xe

\subsubsection{Morphosyntactic operations within the noun phrase}

It has been pointed out above that nouns encode animacy. This has repercussions
in the choice of case markers of the agent and patient cases, which need to
agree with the lexical head they attach to. An example of this is given in 
(\ref{ex:animcaseagr}).

\pex\label{ex:animcaseagr}
\a\label{ex:animok} %
\begin{forest}
[\anno{\xhead{N}}
	[\anno{N\tsub{stem}}
		[{%
			gan \\
			\ups{\Anim} = $+$ \\
		}]
	]
	[\anno{-N\tsub{infl}}
		[{%
			-ang \\
			\ups{\Anim} = $+$ \\
			\ups{\Case} = \Aarg{} \\
		}]
	]
]
\end{forest}

\a\label{ex:animclash} %
\ljudge*\begin{forest}
[\anno{\xhead{N}}
	[\anno{N\tsub{stem}}
		[{%
			gan \\
			\ups{\Anim} = $+$ \\
		}]
	]
	[\anno{N\tsub{infl}}
		[{%
			-reng \\
			\ups{\Anim} = $-$ \\
			\ups{\Case} = \Aarg{} \\
		}]
	]
]
\end{forest}
\xe

Example (\ref{ex:animok}) shows a well-formed construction: the noun,
\xayr{gnF}{gan}{child}, is animate, hence the case particle also needs to be 
animate---the case particle must thus be \rayr{/As}{-ang} to be coherent. In
contrast to this, example (\ref{ex:animclash}) is not well-formed in that the
noun is animate but the case particle, \rayr{/reNF}{-reng}, signals that it is
inanimate: the \Anim{} values of the noun stem and its suffix clash and cannot
be conclusively unified for \xhead{N} itself. The same principle of coherence
is, of course, also true for proper nouns, which receive a case-marking
particle:

\pex\label{ex:animcaseagrname}
\a\label{ex:animokname} %
\begin{forest}
[\anno{\xhead{N}}
	[\anno{Cl}
		[{%
			ang \\
			\ups{\Anim} = $+$ \\
			\ups{\Case} = \Aarg{} \\
		}]
	]
	[\anno{\xhead{N}}
		[{%
			Dita \\
			\ups{\Anim} = $+$ \\
		}]
	]
]
\end{forest}

\a\label{ex:animclashname} %
\ljudge*\begin{forest}
[\anno{\xhead{N}}
	[\anno{Cl}
		[{%
			eng \\
			\ups{\Anim} = $-$ \\
			\ups{\Case} = \Aarg{} \\
		}]
	]
	[\anno{\xhead{N}}
		[{%
			Dita \\
			\ups{\Anim} = $+$ \\
		}]
	]
]
\end{forest}
\xe

Furthermore, example (\ref{ex:nounqadj}) already showed that nouns may be
modified by quantifiers, also compare \autoref{subsec:quantifiers}. In these
cases, plural marking is suppressed by the presence of the modifier which
supplies the information by itself so that further morphological plural marking
by the suffix \rayr{/ye}{-ye} on the noun stem itself is redundant, and as such
limited to certain occasions (see \autoref{sec:numerals}, 
p.~\pageref{hundreds}).

...

% \subsection{Interactions between DP and NP}
% \label{subsec:dpnpinteract}

% With common nouns, a determiner which requires a complement (like the
% demonstrative prefixes, for instance) simply cliticizes to the next \xhead{N}
% it commands, which is illustrated in (\ref{ex:det+comn}). If a proper noun is
% involved, however, there need to be two \xhead{D} slots, one for the case
% marker, and another for the demonstrative prefix, which (\ref{ex:det+propn})
% gives an example of.
% %: \xbar{D} is simply doubled to accommodate for another \xhead{D}.
% % FIXME: If this is possible!?
% In both cases, the determiner prefix is unstressed and combines 
% phonetically with the NP modifier, as shown in (\ref{ex:det+comn}). If there is
% no coalescence of vowels, as in (\ref{ex:det+propn}), the prefix remains
% unstressed: \rayr{Ad/YnF}{ada-Yan} [a.da.ˈjan].

% \begin{multicols}{2}
% \pex\label{ex:det+comn}
% \a\begingl
% 	\gla ayonang //
% 	\glb ayon-ang //
% 	\glc man-\Aarg{} //
% 	\glft `the/a man' //
% \endgl

% \a\begingl
% 	\gla adāyonang //
% 	\glb ada=ayon-ang //
% 	\glc that=man-\Aarg{} //
% 	\glft `that man' //
% \endgl

% \xe

% \pex~\label{ex:det+propn}
% \a\begingl
% 	\gla {ang Diyan} //
% 	\glb ang=Diyan //
% 	\glc \Aarg{}=Diyan //
% 	\glft `Diyan' //
% \endgl

% \a\label{ex:case+det+propn}\begingl
% 	\gla {ang ada-Diyan} //
% 	\glb ang=ada=Diyan //
% 	\glc \Aarg{}=that=Diyan //
% 	\glft `that Diyan' //
% \endgl

% \xe

% \end{multicols}

% \begin{multicols}{2}
% \pex~
% \a \begin{minipage}[t]{\linewidth}
% C-structure for (\ref{ex:det+comn}):\medskip

% \quad {\begin{forest} shorter edges, baseline,
% [NP
% 	[\xbar{N}
% 		[\xhead{D}
% 			[ada-]
% 		]
% 		[\xbar{N}
% 			[\xhead{N}
% 				[ayonang]
% 			]
% 		]
% 	]
% ]
% \end{forest}
% }
% \end{minipage}

% \a \begin{minipage}[t]{\linewidth}
% C-structure for (\ref{ex:det+propn}):\medskip

% \quad {\begin{forest} shorter edges, baseline,
% % [NP
% % 	[\xbar{N}
% % 		[\xhead{D}
% % 			[ang]
% % 		]
% % 		[\xbar{N}
% % 			[\xhead{D}
% % 				[ada-]
% % 			]
% % 			[\xbar{N}
% % 				[\xhead{N}
% % 					[Diyan]
% % 				]
% % 			]
% % 		]
% % 	]
% % ]
% [...]
% \end{forest}
% }
% \end{minipage}
% \xe
% \end{multicols}

% Furthermore, as mentioned before in \autoref{subsubsec:verbprefixes}, it is 
% possible to abbreviate \xayr{dnY}{danya}{such one} + adjective as 
% \rayr{d/}{da-} + adjective. This affixing of the indefinite demonstrative 
% essentially derives a complex anaphora. The trees charted in (\ref{ex:adjpron}) 
% illustrate pronominalized adjectives and possessive pronouns.

% \ex\label{ex:adjpron}\labels%
% \parbox[t]{.4\linewidth}{%
% \tl\quad\begingl%
% 	\gla da-kivo //
% 	\glb da=kivo //
% 	\glc one=small //
% 	\glft `the small one' //
% \endgl%
% }%\medskip
% %
% \begin{tabular}[t]{@{} l @{\quad→\quad} l}
% \begin{forest} shorter edges, narrower nodes,
% [DP
% 	[\anno{\xbar{D}}
% 		[\anno{\xhead{D}}
% 			[danya]
% 		]
% 		[{\anno[\pass{\Adj}]{AP}}
% 			[\anno{\xhead{A}}
% 				[kivo]
% 			]
% 		]
% 	]
% ]
% \end{forest}

% &

% \begin{forest} shorter edges, narrower nodes,
% % [AP
% % 	[\anno{\xbar{A}}
% % 		[\anno{\xhead{D}}
% % 			[da-]
% % 		]
% % 		[\anno{\xbar{A}}
% % 			[\anno{\xhead{A}}
% % 				[kivo]
% % 			]
% % 		]
% % 	]
% % ]
% [...]
% \end{forest}
% \end{tabular}

% \parbox[t]{.4\linewidth}{%
% \tl\quad\label{ex:posspron}%
% \begingl%
% 	\gla da-vana //
% 	\glb da=vana //
% 	\glc such=\Second{}.\Gen{} //
% 	\glft `yours' //
% \endgl}%\medskip
% %
% \begin{tabular}[t]{@{} l @{\quad→\quad} l}
% \begin{forest} shorter edges, narrower nodes,
% [DP
% 	[\anno{\xbar{D}}
% 		[\anno{\xhead{D}}
% 			[danya]
% 		]
% 		[{\anno[\pass{\Adj}]{DP}}
% 			[\anno{\xhead{D}}
% 				[vana]
% 			]
% 		]
% 	]
% ]
% \end{forest}

% &

% \begin{forest} shorter edges, narrower nodes,
% % [DP
% % 	[\anno{\xbar{D}}
% % 		[\anno{\xhead{D}}
% % 			[da-]
% % 		]
% % 		[{\anno{\xbar{D}}}
% % 			[\anno{\xhead{D}}
% % 				[vana]
% % 			]
% % 		]
% % 	]
% % ]
% [...]
% \end{forest}
% \end{tabular}
% \xe

% \rayr{d/}{da-} in this case looks like the indefinite demonstrative prefix,
% defined in (\ref{ex:dmorphlex-dem}) above. However, \rayr{d/} in this case does
% not have a strictly deictic meaning, but an anaphoric one---together with the
% adjunct it forms a pronoun of sorts, though unlike \rayr{dnY}{danya} and like
% the indefinite demonstrative, it cannot stand alone, but needs another word to
% lean on as a clitic. Thus, we have to distinguish between the demonstrative
% prefix \rayr{d/}{da-} and the anaphoric prefix \rayr{d/}{da-}.
% \autoref{fig:da-s} compares their morpholexical features.

% \begin{figure}[t]\centering
% \caption{Comparison between two functions of \rayr{d/}{da-}}
% \begin{tabu} to \linewidth {X[1] X[12] X[1] X[12]}
% \toprule\tableheaderfont
% \multicolumn{2}{c}{Demonstrative \rayr{d/}{da-}}
% & \multicolumn{2}{c}{Anaphoric \rayr{d/}{da-}}
% \\
% \toprule
% D
% & \begin{tabular}[t]{l l l}
% 	\ups{\Pers} & \req{} & 3 \\
% 	\ups{\Def} & = & $-$ \\
% 	\ups{\Spec} & = & $-$ \\
% \end{tabular}

% &

% D
% & \begin{tabular}[t]{l l l}
% 	\ups{\Pred} & = & `pro' \\
% 	\ups{\Pers} & = & 3 \\
% 	\ups{\Def} & = & $+$ \\
% 	\ups{\Index{} \Pers} & = & 3 \\
% 	\ups{\Index{} \Num} & = & \{\Sg{}, \Pl\} \\
% 	\ups{\Index{} \Gend} & = & \{\M{}, \F{}, \N{}\} \\
% 	\ups{\Index{} \Anim} & = & $\pm$ \\
% \end{tabular}
% \\

% \bottomrule

% \end{tabu}
% \label{fig:da-s}
% \end{figure}

% The most striking difference in features should be that anaphoric 
% \rayr{d/}{da-} lists a predicator which designates it as a pro-form, whereas
% demonstrative \rayr{d/}{da-} is, essentially, a functional morpheme only, 
% without lexical content. As a pro-form, the anaphoric prefix is able to
% bind another NP, while demonstrative prefix cannot. Furthermore, the 
% demonstrative prefix requires the presence of a third-person term, while the
% anaphoric prefix defines one. Also, the definiteness and specificity values 
% differs in that the demonstrative prefix with its meaning `such a' refers to
% one, but no particular item of a set of alike items, while the anaphoric prefix 
% refers to a certain item, that is, the referent it binds.

% As shown in (\ref{ex:adjpron}), the host of the anaphoric \rayr{d/}{da-} clitic
% may be an adjective or a possessive pronoun. In contrast to NPs, which are
% complements of \xhead{D}, their maximal projections (AP and DP, respectively)
% are adjuncts and thus need to be annotated with `\pass{\Adj}' rather than 
% `\updown{}'. Complex anaphora with anaphoric \rayr{d/}{da-} may also receive
% case marking like any other nominal, whether DP or NP. The variable person
% features---\Num{}, \Gend{}, and \Anim{}---are inherited from the bound NP, so
% that the respective \Aarg{} and \Parg{} case markers are selected in accordance
% with the referred-to NP's \Anim{} feature. Verbs agree in number, gender, and
% animacy accordingly as well if relevant.

% \pex
% \a\begingl
% 	\gla Le ming eryavāng {\normalfont [\tsup{NP}} @ limu nā @ 
% 			{\normalfont ]\tsub{i}}. //
% 	\glb Le ming ery=vāng {} limu-Ø nā {} //
% 	\glc \PatTI{} can use=\Second{}.\Aarg{} {} shirt-\Top{} {} //
% 	\glft `You can use my shirt.' //
% \endgl

% \a\begingl
% 	\gla Da-nāreng\tsub{i} apitu. //
% 	\glb Da=nā-reng apitu //
% 	\glc one=\Fsg{}.\Gen{}-\AargI{} clean //
% 	\glft `Mine is clean.' //
% \endgl

% \xe

% Note, however, that in predicative statements, if a possessive pronoun is in the
% second position, it can take \rayr{d/}{da-}, but it acts similarly to an
% adjective. This means that in this exceptional case, there is no case marking
% for agent or patient, which would otherwise be required of NPs. This is the 
% reason why in \autoref{fig:da-s}, the \Pers{} and \Def{} values are in 
% parenthesis---these features are not encoded by adjectives.

% \pex
% \a\begingl
% 	\gla Adanyāng {\normalfont [\tsup{NP}} @ nangās nā @ 
% 		{\normalfont ]\tsub{j}}. //
% 	\glb Adanya-ang {} nanga-as nā {} //
% 	\glc that-\Aarg{} {} house-\Parg{} \Fsg{}.\Gen{} {} //
% 	\glft `That is my house'. //
% \endgl

% \a\begingl
% 	\gla Adanyāng da-nā\tsub{j}. //
% 	\glb Adanya-ang da=nā //
% 	\glc that-\Aarg{} one=\Fsg{}.\Gen{} //
% 	\glft `It is mine.' //
% \endgl

% \a\begingl
% 	\gla Apitu da-nā\tsub{i}. //
% 	\glb Apitu da=nā //
% 	\glc clean one=\Fsg{}.\Gen{} //
% 	\glft `Mine is clean.' //
% \endgl

% \xe

% While it is possible to derive complex anaphora from adjectives and possessive
% pronouns this way, the process as described above only works for single 
% words, but not for complex phrases. Thus, for instance, the following examples
% are unacceptable:

% \pex
% \a\label{ex:suchadjp}\ljudge*\begingl
% 	\gla Tahayang {\normalfont [\tsup{DP}} @ da- @ {\normalfont 
% 		[\tsup{AP}} @ dano nay kema @ {\normalfont]]}. //
% 	\glb Taha=yang {} da= {} dano nay kema {} //
% 	\glc have=\Fsg{}.\Aarg{} {} one= {} green and striped {} //
% 	\glft `I have a/the green-and-striped one.' //
% \endgl

% \a\label{ex:suchrelc}\ljudge*\begingl
% 	\gla Vāng {\normalfont [\tsup{DP}} @ da- @ {\normalfont[\tsup{CP}} 
% 		@ si ya silvasayang ren @ {\normalfont ]]}. //
% 	\glb Vāng {} da= {} si ya silv-asa=yang ren-Ø {} //
% 	\glc \Second.\Aarg{} {} one= {} \Rel{} \LocT{} 
% 		see-\Hab{}=\Fsg{}.\Aarg{} market-\Top{} {} //
% 	\glft `You are the who-I-always-see-at-the-market one.' //
% \endgl
% \xe

% Example (\ref{ex:suchadjp}) shows an AP consisting of two coordinated 
% adjectives which are supposed to be nominalized here. The DP would normally 
% have to be a patient, since \xayr{th/}{taha-}{have} requires two 
% arguments---a possessor and a possessee, which are conventionally mapped to 
% \Sbj{}/\Aarg{} and \Obj{}/\Parg{}, respectively. We have seen, however, that 
% Ayeri allows nominal phrases unmarked for case only under very special 
% circumstances. The question is, thus, whether a complex phrase can be 
% nominalized as a whole in the first place. The next questions are, then, where 
% the case marker should go:

% \begin{enumerate}[noitemsep]
% 	\item only on the first head,
% 	\item on each conjunct (would \rayr{d/}{da-} also need to be 
% 		distributed, then?),
% 	\item or, essentially, as a clitic at the very end?
% \end{enumerate}

% \noindent If marking went on each conjunct, this would coordinate two APs 
% instead of two \xhead{A}, however: `the green one and the striped one', which
% is not necessarily the same as `the green and striped one'. With the relative
% clause in (\ref{ex:suchrelc}) only suggestions 1 and 3 are possible, with an
% awkward result as well. The solution in either case is to remember that 
% \rayr{d/}{da-} + non-noun is an abbreviation of the full indefinite 
% demonstrative, which basically behaves like a noun in terms of adjuncts:

% \pex
% \a\label{ex:danyaadjp}\begingl
% 	\gla Le tahayang {\normalfont [\tsup{DP}} @ danya {\normalfont 
% 		[\tsup{AP}} @ dano nay kema @ {\normalfont]]}. //
% 	\glb Le taha=yang {} danya-Ø {} dano nay kema {} //
% 	\glc \PatTI{} have=\Fsg{}.\Aarg{} {} one-\Top{} {} green and striped {} //
% 	\glft `I have the green and striped one.' //
% \endgl

% \a\begingl
% 	\gla Vāng {\normalfont [\tsup{DP}} @ danyās {\normalfont[\tsup{CP}} 
% 		@ si ya silvasayang ren @ {\normalfont ]]}. //
% 	\glb Vāng {} danya-as {} si ya silv-asa=yang ren-Ø {} //
% 	\glc \Second.\Aarg{} {} one-\Parg{} {} \Rel{} \LocT{} 
% 		see-\Hab{}=\Fsg{}.\Aarg{} market-\Top{} {} //
% 	\glft `You are the one who I always see at the market.' //
% \endgl
% \xe

% Instead of the construction in (\ref{ex:danyaadjp}) it is also possible to
% pronominalize the first adjective and to include the second one in a relative 
% clause, or to turn the second adjective into a noun adjunct if there is an
% equivalent noun. Since \xayr{kem}{kema}{striped} is derived from the noun
% \xayr{kem}{kema}{stripe}, this strategy is possible here.

% \pex
% \a\begingl
% 	\gla da-danoley si kema //
% 	\glb da=dano-ley si kema //
% 	\glb one=green-\PargI{} \Rel{} striped //
% 	\glft `the green one which is striped' //
% \endgl

% \a\begingl
% 	\gla da-danoley kemayeri //
% 	\glb da=dano-ley kema-ye-ri //
% 	\glc one=green-\PargI{} stripe-\Pl{}-\Ins{} //
% 	\glft `the green one with stripes' //
% \endgl

% \xe

% \subsection{Determiner phrases}
% \label{subsec:dps}

% The functional heads of DPs are formed by determiners. The c-structure of a DP
% corresponding to the linear description in (\ref{ex:dpdef}) is depicted in
% (\ref{ex:dpcstruct}); annotations about feature inheritance for
% c-to-f-structure conversion have been added. If the DP has a dependent phrase
% providing lexical content, it is in the form of an NP complement, whose head
% forms a lexical co-head of \xhead{D}. This NP must be annotated with
% `\updown{}' to signify that its features are inherited to DP wholesale.
% Alternatively, a DP containing a pronoun may also be modified by an NP adjunct,
% a possessive NP, an AP or a relative-clause CP, which all need to be annotated
% with `\pass{\Adj}', since they are not complements.

% \ex\label{ex:dpcstruct}
% \begin{forest}
% [{\anno[\pass{df} \logor{} \pass{gf}]{DP}}
% 	[\anno{\xbar{D}}
% 		[\anno{\xhead{D}}]
% 		[{$\left(\anno[{%
% 		\updown{}\\%
% 		%\logor{} \pass{\XCompl}\\%
% 		\logor{} \pass{\Adj}%
% 	}]{XP}\right)$}]
% 	]
% ]
% \end{forest}
% \xe

% Ayeri does not possess articles as such, though it does have a variety of
% morphemes which occur as clitic prefixes on nouns which can be analyzed to
% fulfill this function. These clitics are the demonstrative prefixes
% \xayr{Ed/}{eda-}{this}, \xayr{Ad/}{ada-}{that}, and \xayr{d/}{da-}{such (a)},
% as well as the inspecificity prefix \xayr{me/}{mə-}{some} (compare
% \autoref{subsec:nounpref}). These prefixes can only be used with nouns, but not
% with personal pronouns (compare \autoref{subsec:perspro}), since personal
% pronouns as well have the status of determiners. Pronouns, as pro-forms, are in
% complementary distribution with NPs containing a noun. Due to affix-order
% considerations concerning the demonstrative prefixes, it also seems advisable
% to analyze the case markers of proper nouns as determiners. In order to capture
% the fact that an NP is facultative with the demonstrative and inspecificity
% prefixes while it must not occur with pronouns, `NP' appears in parentheses
% since a phrasal complement it is not always present. We can define the
% following morpholexic rules for determiners (\xhead{D}):

% % FIXME: For some of these I'm not sure *at all* whether they're adjuncts or
% % whether they should better be classified as complements :O

% \pex %
% % \a\label{ex:dmorphlex-dem}Demonstrative prefixes:\medskip
% % 
% % 	\begin{tabu} to \linewidth {@{} X[15l] X[5l] X[80l]}
% % 	\savetabu{morphlex}
% % 	\rayr{Ed/}{eda-} (`this')
% % 		& D
% % 		& \begin{tabular}[t]{l l l}
% % % 			\ups{\Pred} & = & `this' \\
% % 			\ups{\Pers} & \req{} & 3 \\
% % 			\ups{\Prox} & = & $+$ \\
% % 			\ups{\Dist} & = & $-$ \\
% % 			\ups{\Def} & = & $+$ \\
% % 			\ups{\Spec} & = & $+$ \\
% % 		\end{tabular}
% % 	\end{tabu}\medskip
% % 
% % 	\begin{tabu} {\usetabu{morphlex}}
% % 	\rayr{Ad/}{ada-} (`that')
% % 		& D
% % 		& \begin{tabular}[t]{l l l}
% % % 			\ups{\Pred} & = & `that' \\
% % 			\ups{\Pers} & \req{} & 3 \\
% % 			\ups{\Prox} & = & $-$ \\
% % 			\ups{\Dist} & = & $+$ \\
% % 			\ups{\Def} & = & $+$ \\
% % 			\ups{\Spec} & = & $+$ \\
% % 		\end{tabular}
% % 	\end{tabu}\medskip
% % 	
% % 	\begin{tabu} {\usetabu{morphlex}}
% % 	\rayr{d/}{da-} \mbox{(`such a')}
% % 		& D
% % 		& \begin{tabular}[t]{l l l}
% % % 			\ups{\Pred} & = & `such' \\
% % 			\ups{\Pers} & \req{} & 3 \\
% % 			\ups{\Def} & = & $-$ \\
% % 			\ups{\Spec} & = & $-$ \\
% % 		\end{tabular}
% % 	\end{tabu}
% % 	
% % \needspace{3\baselineskip}
% % \a Inspecificity prefix:\medskip
% % 
% % 	\begin{tabu} {\usetabu{morphlex}}
% % 	\rayr{me/}{mə-} (`whichever')
% % 		& D
% % 		& \begin{tabular}[t]{l l l}
% % 			\ups{\Def} & = & $-$ \\
% % 			\ups{\Spec} & = & $-$ \\
% % 			\ups{\Pers} & \req{} & 3 \\
% % 		\end{tabular}
% % 	\end{tabu}
% % 
% % \needspace{3\baselineskip}
% \a Personal pronouns:\medskip

% 	\begin{tabu} {\usetabu{morphlex}}
% 	(various)
% 		& N
% 		& \begin{tabular}[t]{l l l}
% 			\ups{\Pred} & = & `pro' \\
% 			\ups{\Pers} & = & \{\First{}, \Second{}, \Third{}\} \\
% 			\ups{\Refl} & = & $\pm$ \\
% 			\ups{\Num} & = & \{\Sg{}, \Pl{}\} \\
% 			\ups{\Gend} & = & \{\M{}, \F{}, \N{}\} \\
% 			\ups{\Anim} & = & $\pm$ \\
% 			\ups{\Case} & = & \{\Aarg{}, \Parg{}, \Dat{}, \Gen{}, 
% 				\Loc{}, \Ins{}, \Caus{}\} \\
% 			\ups{\Def} & = & $+$ \\
% 			\ups{\Spec} & = & $+$ \\
% 		\end{tabular}
% 	\end{tabu}
	
% \needspace{3\baselineskip}
% \a Demonstrative pronouns:\medskip

% 	\begin{tabu} {\usetabu{morphlex}}
% 	(various)
% 		& N
% 		& \begin{tabular}[t]{l l l}
% 			\ups{\Pred} & = & `pro' \\
% 			\ups{\Pers} & = & 3 \\
% 			\ups{\Prox} & = & $\pm$ \\
% 			\ups{\Dist} & = & $\pm$ \\
% 			\ups{\Def} & = & $\pm$ \\
% 			\ups{\Spec} & = & $+$ \\
% 			\ups{\Anim} & = & $\pm$ \\
% 			\ups{\Case} & = & \{\Aarg{}, \Parg{}, \Dat{}, \Gen{}, 
% 				\Loc{}, \Ins{}, \Caus{}\} \\
% 		\end{tabular}
% 	\end{tabu}

% \needspace{3\baselineskip}
% \a Indefinite pronouns:\medskip

% 	\begin{tabu} {\usetabu{morphlex}}
% 	(various)
% 		& N
% 		& \begin{tabular}[t]{l l l}
% 			\ups{\Pred} & = & `pro' \\
% 			\ups{\Pers} & = & 3 \\
% 			\ups{\Def} & = & $-$ \\
% 			\ups{\Spec} & = & $-$ \\
% 			\ups{\Anim} & = & $\pm$ \\
% 			\ups{\Case} & = & \{\Aarg{}, \Parg{}, \Dat{}, \Gen{}, 
% 				\Loc{}, \Ins{}, \Caus{}\} \\
% 		\end{tabular}
% 	\end{tabu}
	
% % \needspace{3\baselineskip}
% % \a\label{ex:dmorphlex-propn}Case markers of proper nouns:\medskip	
	
% % 	\begin{tabu} {\usetabu{morphlex}}
% % 	(various)
% % 		& D
% % 		& \begin{tabular}[t]{l l l}
% % % 			\ups{\Pers} & \req{} & \Third \\
% % % A 3rd person must not be *required* if the topic particle is in fact a D head 
% % % in Spec NP, since pronouns as well can serve as topics without being 3rd 
% % % persons!
% % 			\ups{\Anim} & \req{} & $\pm$ \\
% % 			\ups{\Case} & = & \{\Aarg{}, \Parg{}, \Dat{}, \Gen{}, 
% % 				\Loc{}, \Ins{}, \Caus{}\} \\
% % 		\end{tabular}
% % 	\end{tabu}\medskip
	
% \xe

% Demonstrative prefixes, the inspecificity prefix \rayr{me/}{mə-}, as well as
% proper-noun case markers cannot be without an NP complement, which is what
% `\req{}' is supposed to express in the feature specification: the demonstrative
% prefixes require that an element exist which encodes a third-person referent to
% are assumed to encode third person by default.\footnote{A simple `='
% \emph{defines} a value; a subscript `c' indicates that the morpheme
% \emph{requires} that this value be present. Thus `\req{}' expresses a
% \fw{constraining} equation \citep[59--61]{bresnan2016}.} On the other hand, 
% personal and demonstrative pronouns inherently define information on person 
% (\Pers{}). Since there is a great number of both personal and demonstrative 
% pronouns, only the various values they can assume are indicated, without 
% ensuring that the combinations are actually possible (compare 
% \autoref{subsec:perspro}). Personal pronouns, for instance, only distinguish 
% \Inan{} as set against $\{\M{}, \F{}, \N{}\}$, which are subgroups of \An{}. 
% Gender is also not distinguished in all persons, but only in the third.

% \section{Adjective phrases and adverbial phrases}

% Adjectives describe properties of nouns; adverbs describe properties of actions,
% but may also provide additional information about place, time, manner,
% quantity, degree etc. more generally by not only modifying verbs, but also other
% parts of speech, or even whole clauses. The morphology of adjectives has been 
% elaborated in \autoref{sec:adjectives}; different uses of adverbs have been
% described in \autoref{sec:adverbs}. In this section, I will not deal with those
% adverbials consisting of an adpositional phrase, that is, ones which primarily
% provide information on time and place.

% Adjective phrases are headed by an adjective (\xhead{A}), which may either be
% modified by an adverb (\ref{ex:adjadv}), or by an NP complement, for instance 
% (\ref{ex:adjnp}):

% \pex
% \a\label{ex:adjadv}\begingl
% 	\gla viyu \textbf{cuyam} //
% 	\glb viyu cuyam //
% 	\glc proud indeed //
% 	\glft `indeed proud' //
% \endgl

% \a\label{ex:adjnp}\begingl
% 	\gla viyu \textbf{yaneri} \textbf{tan} //
% 	\glb viyu yan-eri tan //
% 	\glc proud son-\Ins{} \TplM{}.\Gen{} //
% 	\glft `proud of their son' //
% \endgl

% \xe

% The phrase structure of an AP thus can be generalized as shown in 
% (\ref{ex:apstruct}), which corresponds to the c-structure illustrated in 
% (\ref{ex:apcstruct}). I am assuming here that adjectives and adverbs share the 
% same structures, hence I will not distinguish between them formally.

% \ex\label{ex:apstruct}
% AP → \xhead{A} $\left(\anno*[{%
% 		\pass{\Adjc{}}\\
% 		\logor{} \pass{\OblqT}% FIXME: Not sure about this
% 	}]{XP}\right)$
% \xe

% \ex~\label{ex:apcstruct}
% \begin{forest}
% [{\anno[\pass{\Adjc} \logor{} \pass{\XCompl}]{AP}}
% 	[\anno{\xbar{A}}
% 		[\anno{\xhead{A}}]
% 		[{$\left(\anno[{%
% 				\pass{\Adjc{}}\\
% 				\logor{} \pass{\OblqT}% FIXME: Not sure about this
% 			}]{XP}\right)$}]
% 	]
% ]
% \end{forest}
% \xe

% Adjectives and adverbs in Ayeri can be interpreted as not inflecting, with even
% the comparative suffixes construed as adverbial clitics meaning `rather' and
% `most'. The only feature property of both adjectives and adverbs would be 
% \Pred{}, then. A more progressive and strictly lexicalist interpretation would
% treat these markers as inflections, however, so that adjectives and adverbs 
% would both additionally have to have a \textsc{compar} feature (compare 
% \cite[Feature Table]{pargram}) taking the values \Comp{} and \Supl{}.\footnote{%
% In this context, \Comp{} is short for `comparative', not `complement'!} Both
% hypotheses should be easily comprehensible, since clitics usually are 
% grammaticalized as inflection. That such a grammaticalization process is in
% progress here is possible, in that the comparison suffixes---even though they
% still exist as separate lexical items, for which reason I am glossing them as
% clitics rather than as inflection---mainly serve a functional purpose here.
% That is, they may be interpreted as being in the process of losing their 
% \Pred{} value in favor of becoming integrated into the language's grammar as
% functional morphemes. The prototype of a table of morpholexic definitions for 
% \xhead{A} accordingly looks like this:

% \needspace{3\baselineskip}
% \ex \begin{tabu}[t] {\usetabu{morphlex}}
% 	...
% 		& A
% 		& \begin{tabular}[t]{l l l}
% 			\ups{\Pred} & = & `...' \\
% 			\ups{\Compar} & = & \{\Comp{}, \Supl{}\} \\
% 		\end{tabular}
% 	\end{tabu}

% \xe

% \section{Adpositional phrases}
%
% \section{Inflectional, verb, and complementizer phrases}
%
% * Ayeri's unmarked word order is essentially VSO, maaaaaaaybe like this???
%
%   IP
%     \
%      I'
%     /  \
%    D°   I'
%        /  \
%       I°   S
%           / \
%          NP XP*
% 
%   → Analysis similar to the one suggested for Irish by ??? and ??? 
%     (Kroeger 1991, 1993; Bresnan 2016)
%   → Probably one has to distinguish clauses with IP from `small clauses' 
% 	  (cf. exposition to Kroeger 1991) with only S
%   → S is likely an exocentric category
% 	  → support/tests for this?
% 	  → Kroeger 1991 should list some criteria in this respect
% * Subject vs. Topic vs. Actor vs. Nominative in Ayeri, in relation to
%   superficial (?) similarities with Austronesian alignment (Kroeger 1991,
%   1993, 2007; Schachter 2015)
%   → Promoting patient in whatever way does not demote logical subject etc.
%   → Ayeri's notion of A = logical subject (also what LFG calls `SUBJ', right?)
%     probably stronger than Tagalog's
%   → PROBLEM: terminology all about the place---may have inspired differences 
%     between Ayeri and real-world AA + my only now beginning to learn more
%     about syntax and stuff
%     → IMPORTANT: how do I use the respective terms and why?
% * Bresnan (2016) on extended head principle relevant w/r/t placement of 
%   finite verb in I°?
%   → Are there any tests for whether there's a covert VP after all? Nothing
%     verbal ever seems to occur in S'
%   → Ayeri should treat logical subjects and objects similarly, so no 
%     hierarchy between them?
% * Interesting problem: subject pronoun citics as a specifier of S?
%   → Funny stuff happens when adverbs are present
%   → Especially with patient-pronoun subjects (if that term is appropriate)
% * Patient subjects, causative constructions, no secundative, avoidance (?) of 
%   raising all need to be investigated with regards to proto-role mapping in 
%   a-structure ([±r, ±o])
% * Existential vs. predicative statements
%   → also, object predicatives!
%   → also, comparative verbs: kama-, eng-, va- and why they're weird
% * Agreement with conjoined/disjoined NPs
%   → due to the way verb agreement in general works, closest-conjunct agreement
%     is the most likely strategy for gender mismatches; yet number resolution.
%   → default gender for anaphoric reference to mixed-gender NPs (also animacy
%     mismatches!)
%
% \citet[153--156, 165]{bresnan2016} → pronouns to clitics to agreement
% \citet[157]{bresnan2016} → evidence criteria for IP
% \citet[151\psqq]{dixon2012} → subjects
% \citet[116\psqq]{comrie1989} → subjects
% \citet[26--54]{kroeger1991} → subjects (and actually the whole thesis)
%
